---
title: "Hop ROB Sensitivity Analysis"
author: "Mick Girdwood"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, error = FALSE)

black <- "#000000"
orange <- "#E69F00"
lightblue <- "#56B4E9"
green <- "#009E73"
yellow <- "#F0E442"
blue <- "#0072B2"
red <- "#D55E00"
pink <- "#CC79A7"

library(tidyverse)

within_data <- read_csv("data/processed/within_data.csv") %>%
  filter(!study %in% c("Geoghegan 2007", "Karanikas 2004", "Karanikas 2005", "Nicholas 2001", 
                       "Urabe 2002", "Tate 2017", "Dalton 2011", "Hall 2015", "Noehren 2014", "Rahova 2020",
                       "Thomas 2013", "Balki 2019", "Rhatomy", "Sullivan 2022"))

# filtering studies that were included in hip review (where no limit on publication year), but need to remove forthis review
casecontrol <- read_csv("data/processed/casecontrol.csv") %>%
  filter(!study %in% c("Geoghegan 2007", "Karanikas 2004", "Karanikas 2005", "Nicholas 2001", 
                       "Urabe 2002", "Tate 2017", "Dalton 2011", "Hall 2015", "Noehren 2014", "Rahova 2020",
                       "Thomas 2013", "Balki 2019", "Rhatomy", "Sullivan 2022"))

rob <- read_csv("data/processed/rob_final.csv") %>%
  mutate(rob_analysis = case_when(
    outcome_c != "Low" & blinding_assessor_c != "Low" & selection_c != "Low" ~ "High",
    TRUE ~ "Low"))

within_data <- left_join(within_data, rob %>% select(study, rob_analysis), by = "study")
casecontrol <- left_join(casecontrol, rob %>% select(study, rob_analysis), by = "study")


## Functions:


pointsfunction <- function(data){
  data %>%
    select(ci.ub, x) %>% 
    bind_rows(., data %>% 
                select(ci.lb, x) %>% 
                rename(ci.ub = ci.lb) %>% 
                arrange(desc(x)))
}

pi_pointsfunction <- function(data){
  data %>%
    select(pi.ub, x) %>% 
    bind_rows(., data %>% 
                select(pi.lb, x) %>% 
                rename(pi.ub = pi.lb) %>% 
                arrange(desc(x)))
}

mv_plotfunction_rob <- function(model, model2){
  
  knot_pos <- NULL # set to 0 for logic to work later
  
  moderator <- as.character(model$call$mods[2]) # get the moderator variable
  data <- model$data # get the data
  
  last_timepoint <- round(max(data$timepoint_mean),0)
  
  total_k  <- model$k
  total_studies <- data %>% filter(!is.na(vi)) %>% summarise(n = length(unique(study))) %>% as.numeric()
  total_n <- data %>% distinct(cohort, .keep_all = TRUE)  %>% filter(!is.na(vi)) %>% summarise(n = sum(acl_n)) %>% as.numeric
  
  if (str_detect(moderator, "rcs")) { # if a rcs is used, extract the information
    spline <- as.list(model$formula.mods[[2]][[3]]) 
    knot_pos <- unlist(spline[sapply(spline, is.numeric)]) # this gets the vector of knot positions or no. of knots
  }
  
  # depending on the type od moderator (i.e. log, linear, rcs...) calculate the predicted curve points
  mod_type <- if (str_detect(moderator, "log")) {
    
    #log
    points <- data.frame(predict(robust(model, cluster = cohort, clubSandwich = TRUE), 
                                 newmods = log(seq(1,100, length = 100)),
                                 transf = exp
    )) %>% mutate(x = row_number())
    
  } else if (str_detect(moderator, "poly")) {
    
    # poly
    points <- data.frame(predict(robust(model, cluster = cohort, clubSandwich = TRUE), 
                                 newmods=unname(poly(seq(1,100, length = 100), degree = 2, raw=TRUE)),
                                 transf = exp
    )) %>% mutate(x = row_number())
    
  } else if (length(knot_pos) == 3) {
    
    # 3 knot rcs
    knots <- attr(rcs(model.matrix(model)[,2], knot_pos), "parms")
    points <- data.frame(predict(robust(model, cluster = cohort, clubSandwich = TRUE), 
                                 newmods=rcspline.eval(seq(1,100, length = 100), knots, inclx=TRUE),
                                 transf = exp)) %>% 
      mutate(x = row_number())
    
  } else if (length(knot_pos) == 1) { # i.e. cases where only the number of knots is given
    
    # 3 knot rcs
    knots <- attr(rcs(data$timepoint_mean, as.numeric(knot_pos)), "parms") # use model data to get knot positions
    points <- data.frame(predict(robust(model, cluster = cohort, clubSandwich = TRUE), 
                                 newmods=rcspline.eval(seq(1,100, length = 100), knots, inclx=TRUE),
                                 transf = exp)) %>% 
      mutate(x = row_number())
    
  } else if (length(knot_pos) == 4) {
    
    # 4 knot rcs
    knots <- attr(rcs(model.matrix(model)[,2], knot_pos), "parms")
    points <- data.frame(predict(robust(model, cluster = cohort, clubSandwich = TRUE), 
                                 newmods=rcspline.eval(seq(1,100, length = 100), knots, inclx=TRUE),
                                 transf = exp)) %>% 
      mutate(x = row_number())
    
  } else {
    # Linear
    points <- data.frame(predict(robust(model, cluster = cohort, clubSandwich = TRUE), 
                                 newmods = seq(1,100, length = 100),
                                 transf = exp)) %>% mutate(x = row_number())
  }
  
  # calculate ci points
  ci <-  pointsfunction(points)
  
  ci <- ci %>% filter(x < last_timepoint + 6)
  points <- points %>% filter(x < last_timepoint + 6)
  
  
  #### 2nd model
  
  knot_pos2 <- NULL # set to 0 for logic to work later
  
  moderator2 <- as.character(model2$call$mods[2]) # get the moderator variable
  data2 <- model2$data # get the data
  
  last_timepoint2 <- round(max(data2$timepoint_mean),0)
  
  if (str_detect(moderator, "rcs")) { # if a rcs is used, extract the information
    spline2 <- as.list(model2$formula.mods[[2]][[3]]) 
    knot_pos2 <- unlist(spline[sapply(spline, is.numeric)]) # this gets the vector of knot positions or no. of knots
  }
  
  # depending on the type od moderator (i.e. log, linear, rcs...) calculate the predicted curve points
  mod_type2 <- if (str_detect(moderator, "log")) {
    
    #log
    points2 <- data.frame(predict(robust(model2, cluster = cohort, clubSandwich = TRUE), 
                                 newmods = log(seq(1,100, length = 100)),
                                 transf = exp
    )) %>% mutate(x = row_number())
    
  } else if (str_detect(moderator, "poly")) {
    
    # poly
    points2 <- data.frame(predict(robust(model2, cluster = cohort, clubSandwich = TRUE), 
                                 newmods=unname(poly(seq(1,100, length = 100), degree = 2, raw=TRUE)),
                                 transf = exp
    )) %>% mutate(x = row_number())
    
  } else if (length(knot_pos) == 3) {
    
    # 3 knot rcs
    knots2 <- attr(rcs(model.matrix(model2)[,2], knot_pos2), "parms")
    points2 <- data.frame(predict(robust(model2, cluster = cohort, clubSandwich = TRUE), 
                                 newmods=rcspline.eval(seq(1,100, length = 100), knots2, inclx=TRUE),
                                 transf = exp)) %>% 
      mutate(x = row_number())
    
  } else if (length(knot_pos2) == 1) { # i.e. cases where only the number of knots is given
    
    # 3 knot rcs
    knots2 <- attr(rcs(data2$timepoint_mean, as.numeric(knot_pos2)), "parms") # use model data to get knot positions
    points2 <- data.frame(predict(robust(model2, cluster = cohort, clubSandwich = TRUE), 
                                 newmods=rcspline.eval(seq(1,100, length = 100), knots2, inclx=TRUE),
                                 transf = exp)) %>% 
      mutate(x = row_number())
    
  } else if (length(knot_pos2) == 4) {
    
    # 4 knot rcs
    knots2 <- attr(rcs(model.matrix(model2)[,2], knot_pos2), "parms")
    points2 <- data.frame(predict(robust(model2, cluster = cohort, clubSandwich = TRUE), 
                                 newmods=rcspline.eval(seq(1,100, length = 100), knots2, inclx=TRUE),
                                 transf = exp)) %>% 
      mutate(x = row_number())
    
  } else {
    # Linear
    points2 <- data.frame(predict(robust(model2, cluster = cohort, clubSandwich = TRUE), 
                                 newmods = seq(1,100, length = 100),
                                 transf = exp)) %>% mutate(x = row_number())
  }
  
  # calculate ci points
  ci2 <-  pointsfunction(points2)
  
  ci2 <- ci2 %>% filter(x < last_timepoint2 + 6)
  points2 <- points2 %>% filter(x < last_timepoint2 + 6)
  
  data3 <- data2 %>% filter(rob_analysis == "Low")
  
  total_k2  <- model2$k
  total_studies2 <- data3 %>% filter(!is.na(vi)) %>% summarise(n = length(unique(study))) %>% as.numeric()
  total_n2 <- data3 %>% distinct(cohort, .keep_all = TRUE)  %>% filter(!is.na(vi)) %>% summarise(n = sum(acl_n)) %>% as.numeric
  
  
  
  total_nrob <- paste0(total_n, " - ", total_n - total_n2)
  rob <- total_studies - total_studies2

    
    # plot data and predicted model
    plot <- data %>%
      ggplot(aes(x = timepoint_mean, y = exp(yi), group = cohort)) +
      geom_point(aes(size = acl_n, colour = rob_analysis), alpha = 0.3) + 
      scale_colour_manual(values = c("red", "black"), name = "ROB") +
      geom_line(alpha = 0.8, colour = "grey") + 
      scale_y_continuous(breaks = c(0.6, 0.8, 1.0, 1.2), labels = c("-40%", "-20%", "0%", "+20%")) +
      #scale_y_continuous(labels = scales::percent, limits = c(-0.75, 0.25)) +
      scale_x_continuous(breaks = c(0, 12, 24, 60, 120), labels = c(0, 1, 2, 5, 10)) +
      coord_cartesian(xlim = c(0,100), ylim = c(0.5, 1.2)) +
      scale_size(range = c(0, 10)) +
      labs(x = "Time since surgery (Months)", y = "Percentage Deficit", colour = "Measure", size = "Participants (n)") +
      #geom_smooth(aes(x = timepoint_mean, y = yi), method = "lm", colour = "green", inherit.aes = FALSE) +
      #geom_abline(intercept = -0.2198, slope = 0.0014, colour = "red") +
      geom_line(data = points, aes(x = x, y = pred), colour = "red", linewidth = 1.3, inherit.aes = FALSE) +
      geom_polygon(data = ci, aes(x = x, y = ci.ub), fill = "red",  alpha = 0.1, inherit.aes = FALSE) +
      geom_line(data = points2, aes(x = x, y = pred), colour = "blue", linewidth = 1.3, inherit.aes = FALSE) +
      geom_polygon(data = ci2, aes(x = x, y = ci.ub), fill = "blue",  alpha = 0.1, inherit.aes = FALSE) +
      annotate(geom = "text", x = 100, y = 0.5, label = paste0(rob, " studies removed"), family = "Karla", hjust = 1) +
      annotate(geom = "text", x = 100, y = 0.55, label = paste0("n = ", total_nrob), family = "Karla", hjust = 1) +
      theme_mgpub() +
            theme(panel.grid.major.x = element_line(linewidth = rel(0.5), linetype = 2))

  
  return(plot)
  
}
```

```{r libraries}
library(metafor) # for meta-analysis
library(clubSandwich) # for covariance matrix and robust ci estimates
library(rms) # for fitting splines
library(ellipse)
library(ggpubr)
library(mgfunctions)
library(kableExtra)
```
# Within-Person

## 1. Single Hop - Within Person

```{r singlehop}

# Single Leg Hop Data
singlehop_data <- within_data %>%
  rename(acl_graft_group = graft_group) %>%
  filter(str_detect(measure, "single hop")) %>%
  filter(timepoint_mean >0.2, # no pre-operative data
         timepoint_mean < 120, # limiting to 10 years post ACLR
         str_detect(graft, "contralateral|Contralateral", negate = TRUE)) %>% #
  mutate(es_id = row_number()) %>%
  filter(study != "Karikis 2016")

```


### Specify covariance matrix

```{r single cov}
hopV <- impute_covariance_matrix(vi = singlehop_data$vi, 
                                  cluster = singlehop_data$cohort, # cluster by cohort
                                  ti = singlehop_data$timepoint_mean, # timepoint variable
                                  ar1 = 0.85, # auto-correlation between timepoints
                                  check_PD = TRUE, # check positive/defininte
                                  smooth_vi = TRUE,
                                  return_list = FALSE) # return the full matrix

```

### Single Leg Hop

```{r singlehop final}

singlehop_mv <- rma.mv(yi, hopV, 
                       mods = ~rcs(timepoint_mean, 4),
                          data = singlehop_data, 
                          random = list(~ timepoint_mean|cohort),
                          struct = c("CAR"))

singlehop_mv_rob <- rma.mv(yi, hopV, 
                       mods = ~rcs(timepoint_mean, 4),
                          data = singlehop_data, 
                          random = list(~ timepoint_mean|cohort),
                          struct = c("CAR"),
                       subset = rob_analysis == "Low")

mv_plotfunction_rob(singlehop_mv, singlehop_mv_rob)


```

## 2. Triple Hop

```{r triplehop}
triplehop_data <- within_data %>%
  rename(acl_graft_group = graft_group) %>%
  filter(str_detect(measure, "triple hop")) %>%
  filter(timepoint_mean >2, # no pre-operative data
         timepoint_mean < 120,
         str_detect(graft, "contralateral|Contralateral", negate = TRUE)) %>% 
  mutate(es_id = row_number())

```

### Specify covariance matrix

```{r triple cov}

triplehopV <- impute_covariance_matrix(vi = triplehop_data$vi, 
                                 cluster = triplehop_data$cohort,
                                 ti = triplehop_data$timepoint_mean,
                                 ar1 = 0.85,
                                 check_PD = TRUE,
                                 smooth_vi = TRUE,
                                 return_list = FALSE)

```

### Triple Hop 
(not crossover)

```{r triplehop final}

triplehop_mv <- rma.mv(yi, triplehopV, 
                       mods = ~log(timepoint_mean),
                       data = triplehop_data, 
                       random = list(~ timepoint_mean|cohort),
                       struct = c("CAR"))

triplehop_mv_rob <- rma.mv(yi, triplehopV, 
                       mods = ~log(timepoint_mean),
                       data = triplehop_data, 
                       random = list(~ timepoint_mean|cohort),
                       struct = c("CAR"),
                       subset = rob_analysis == "Low")

mv_plotfunction_rob(triplehop_mv, triplehop_mv_rob)


```

## 3. Triple Crossover Hop

```{r triplexhop}
triplexhop_data <- within_data %>%
  rename(acl_graft_group = graft_group) %>%
  filter(str_detect(measure, "triple crossover hop")) %>%
  filter(timepoint_mean >2, # no pre-operative data
         timepoint_mean < 120,
         str_detect(graft, "contralateral|Contralateral", negate = TRUE)) %>% 
  mutate(es_id = row_number()) 

```

### Specify covariance matrix

```{r triplex cov}
triplexhopV <- impute_covariance_matrix(vi = triplexhop_data$vi, 
                                       cluster = triplexhop_data$cohort,
                                       ti = triplexhop_data$timepoint_mean,
                                       ar1 = 0.85,
                                       check_PD = TRUE,
                                       smooth_vi = TRUE,
                                       return_list = FALSE)
```

### Final Model - Triple Crossover Hop 

```{r triplexhop final}

triplexhop_mv <- rma.mv(yi, triplexhopV, 
                       mods = ~log(timepoint_mean),
                       data = triplexhop_data, 
                       random = list(~ timepoint_mean|cohort),
                       struct = c("CAR"))

triplexhop_mv_rob <- rma.mv(yi, triplexhopV, 
                       mods = ~log(timepoint_mean),
                       data = triplexhop_data, 
                       random = list(~ timepoint_mean|cohort),
                       struct = c("CAR"),
                       subset = rob_analysis == "Low")

mv_plotfunction_rob(triplexhop_mv, triplexhop_mv_rob)

```

## 4. Side Hop

```{r sidehop}
sidehop_data <- within_data %>%
  rename(acl_graft_group = graft_group) %>%
  filter(str_detect(measure, "side hop")) %>%
  filter(timepoint_mean >2, # no pre-operative data
         timepoint_mean < 120,
         str_detect(graft, "contralateral|Contralateral", negate = TRUE)) %>% 
  mutate(es_id = row_number()) 

```

### Specify covariance matrix

```{r sidehop cov}
sidehopV <- impute_covariance_matrix(vi = sidehop_data$vi, 
                                        cluster = sidehop_data$cohort,
                                        ti = sidehop_data$timepoint_mean,
                                        ar1 = 0.85,
                                        check_PD = TRUE,
                                        smooth_vi = TRUE,
                                        return_list = FALSE)
```

### Final Model - Side Hop 

```{r sidehop final}

sidehop_mv <- rma.mv(yi, sidehopV, 
                        mods = ~rcs(timepoint_mean, 3),
                        data = sidehop_data, 
                        random = list(~ timepoint_mean|cohort),
                        struct = c("CAR"))

sidehop_mv_rob <- rma.mv(yi, sidehopV, 
                        mods = ~rcs(timepoint_mean, 3),
                        data = sidehop_data, 
                        random = list(~ timepoint_mean|cohort),
                        struct = c("CAR"),
                        subset = rob_analysis == "Low")

mv_plotfunction_rob(sidehop_mv, sidehop_mv_rob)

```

## 5. Vertical Hop

```{r verticalhop}
verticalhop_data <- within_data %>%
  rename(acl_graft_group = graft_group) %>%
  filter(str_detect(measure, "vertical hop")) %>%
  filter(timepoint_mean >2, # no pre-operative data
         timepoint_mean < 120,
         str_detect(graft, "contralateral|Contralateral", negate = TRUE)) %>% 
  mutate(es_id = row_number()) 
```

### Specify covariance matrix

```{r verticalhop cov}
verticalhopV <- impute_covariance_matrix(vi = verticalhop_data$vi, 
                                     cluster = verticalhop_data$cohort,
                                     ti = verticalhop_data$timepoint_mean,
                                     ar1 = 0.85,
                                     check_PD = TRUE,
                                     smooth_vi = TRUE,
                                     return_list = FALSE)
```

### Final Model - Vertical Hop 

```{r verticalhop final}

verticalhop_mv <- rma.mv(yi, verticalhopV, 
                     mods = ~log(timepoint_mean),
                     data = verticalhop_data, 
                     random = list(~ timepoint_mean|cohort),
                     struct = c("CAR"))

verticalhop_mv_rob <- rma.mv(yi, verticalhopV, 
                     mods = ~log(timepoint_mean),
                     data = verticalhop_data, 
                     random = list(~ timepoint_mean|cohort),
                     struct = c("CAR"),
                      subset = rob_analysis == "Low")

mv_plotfunction_rob(verticalhop_mv, verticalhop_mv_rob) + scale_colour_manual(values = c("black"), name = "ROB")

```


## 6. 6m timed Hop

```{r sixmhop}

sixmhop_data_pre <- within_data %>%
  rename(acl_graft_group = graft_group) %>%
  filter(str_detect(measure, "6m timed hop")) %>%
  filter(timepoint_mean >2, # no pre-operative data
         timepoint_mean < 120,
         str_detect(graft, "contralateral|Contralateral", negate = TRUE)) %>% 
  mutate(es_id = row_number()) 

# Some studies invert 6m hop data results (as 6m hop score = better = lower score)
# Have individually checked all studies, some need to be switched

# These studies report the raw for injured mean and non-injured, so to calcualte LSI, need to take non-inj / inj
invert6m <- sixmhop_data_pre %>%
  filter(study %in% c("Ebert 2018", "Norte 2020", "Ebert 2019", "Meierbachtol 2017", "Logerstedt 2013", "Moran 2022a",
                      "Sonesson 2021b", "Sonesson 2021a")) %>%
  escalc(ni = acl_n, m1i = noninj_mean, sd1i = noninj_sd, m2i = inj_mean, sd2i = inj_sd, ri = ri, 
         data = ., measure = "ROMC") %>%
  mutate(yi = case_when(
           is.na(yi) & !is.na(lsi_mean) ~ log(lsi_mean),
           TRUE ~ yi),
         vi = case_when(
           is.na(vi) & !is.na(lsi_mean) ~ lsi_sd^2/(acl_n * lsi_mean^2),
           TRUE ~ vi))

# These studies only report LSI, take the inverse of the LSI
invert6mlsi <- sixmhop_data_pre %>%
  filter(study %in% c("Norte 2019b", "Arhos 2020a", "Barnett 2020", "Reinke 2011")) %>%
  select(-c(yi:ci.ub)) %>%
  mutate(lsi_mean = 1/lsi_mean) %>%
  mutate(yi = log(lsi_mean),
         vi = lsi_sd^2/(acl_n * lsi_mean^2))

# Join back Together
sixmhop_data <- bind_rows(sixmhop_data_pre %>% filter(!study %in% c("Ebert 2018", "Norte 2020", "Ebert 2019", "Meierbachtol 2017", "Logerstedt 2013", "Moran 2022a",
                                                "Sonesson 2021b", "Sonesson 2021a", "Norte 2019b", "Arhos 2020a", "Barnett 2020", "Reinke 2011")),
          invert6m) %>%
  bind_rows(., invert6mlsi) 
```

### Specify covariance matrix

```{r sixmhop cov}

sixmhopV <- impute_covariance_matrix(vi = sixmhop_data$vi, 
                                         cluster = sixmhop_data$cohort,
                                         ti = sixmhop_data$timepoint_mean,
                                         ar1 = 0.85,
                                         check_PD = TRUE,
                                         smooth_vi = TRUE,
                                         return_list = FALSE)
```

### Final Model - 6m timed hop 

```{r sixmhop final}

sixmhop_mv <- rma.mv(yi, sixmhopV, 
                         mods = ~log(timepoint_mean),
                         data = sixmhop_data, 
                         random = list(~ timepoint_mean|cohort),
                         struct = c("CAR"))

sixmhop_mv_rob <- rma.mv(yi, sixmhopV, 
                         mods = ~log(timepoint_mean),
                         data = sixmhop_data, 
                         random = list(~ timepoint_mean|cohort),
                         struct = c("CAR"),
                         subset = rob_analysis == "Low")

mv_plotfunction_rob(sixmhop_mv, sixmhop_mv_rob)


```




# Case Control

Significantly less data available for between person/case control comparisons. Only single hop has enough data to run same analysis as within person.

## 1. Single Hop - Between Person

```{r singlehopcont}

singlehopcont_data <- casecontrol %>%
  rename(timepoint_mean = acl_timepoint_mean) %>%
  filter(str_detect(measure, "single hop")) %>%
  filter(timepoint_mean >2, # no pre-operative data
         timepoint_mean < 120) %>% 
  mutate(es_id = row_number()) 

```


### Specify covariance matrix


```{r singlehopcont cov}
hopcontV <- impute_covariance_matrix(vi = singlehopcont_data$vi, 
                                 cluster = singlehopcont_data$cohort,
                                 ti = singlehopcont_data$timepoint_mean,
                                 ar1 = 0.85,
                                 check_PD = TRUE,
                                 smooth_vi = TRUE,
                                 return_list = FALSE)

```

### Final Model - Single Leg Hop

```{r singlehopcont final}

singlehopcont_mv <- rma.mv(yi, hopcontV, 
                       mods = ~log(timepoint_mean),
                       data = singlehopcont_data, 
                       random = list(~ timepoint_mean|cohort),
                       struct = c("CAR"))

singlehopcont_mv_rob <- rma.mv(yi, hopcontV, 
                       mods = ~log(timepoint_mean),
                       data = singlehopcont_data, 
                       random = list(~ timepoint_mean|cohort),
                       struct = c("CAR"),
                       subset = rob_analysis == "Low")

mv_plotfunction_rob(singlehopcont_mv, singlehopcont_mv_rob)

```


